---
title: "sandbox-model-sim"
author: "Curtis Kephart"
date: "2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE, message=F}
require(ggplot2)
require(dplyr)
require(ggvis)
require(shiny)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```


****

## Tobler's hiking function 

Walking velocity:
$$
W=6e^{\displaystyle-3.5\left\vert\frac{dh}{dx}+0.05\right\vert}
$$

$$
\frac{dh}{dx}=S=\tan\Theta
$$

where

- $dh$ = elevation difference,
- $dx$ = distance,
- $S$ = slope,
- $\Theta$ = angle of slope (inclination).


```{r base-tobler, echo = F, warning=F}
slopes <- seq(-.80,.80, by = 0.01)
walkspeed_1 <- function(slope = -0.05, 
                        maxS = 3.4, # max speed
                        bSlope = 0.05 # fastest slope
                        ){
  maxS * exp(-3.5 * abs(tan(slope) + bSlope))
}
speeds <- walkspeed_1(slopes)


ggplot(
  data = (
    data.frame(
      slope = slopes,
      speed = speeds
    )
  ),
  aes(
    x = slope,
    y = speed
  )
) + 
  geom_line() +
  theme_minimal() +
  xlab("Slope (in degrees)") +
  ylab("Walking Speed (mph)") +
  ggtitle("Tobler's hiking function")

```





****

## Modified Tobler's  Trail Speed Model
Walking velocity:
$$
V = V_m e^{\displaystyle(Sp)\left\vert\frac{dh}{dx}+0.05\right\vert}
$$

$$
\frac{dh}{dx} = S = \tan\Theta
$$

where

- $S$ = slope,
- $dh$ = elevation difference,
- $dx$ = distance,
- $V_m$ = maximum speed.
- $\phi$ = fitness
- $Sp$ = shape of relationships between slope and speed
- $\Theta$ = angle of slope (inclination).

The model continues to assume top travel speeds are at a downward angle of 5 degrees. 

A slope beyond 90 degrees means you are either rock climbing or falling, and so velocity approximately zero. 

Given sufficient hiker speed observations over various elevation profiles and different experience levels, we can estimate the model's maximum speed, fitness, and shape parameters. 




```{r mod-tobler, echo=F, cache = F, fig.width=6}

walkspeed <- function(slope = -0.05, 
                      V_m = 3.4, # max speed
                      Sp = -3.5, # shape
                      bSlope = 0.05, # fastest slope
                      phi = 1.5, # fitness
                      week = 1 # of hiking
                      ){
  (V_m - phi/week) * exp((Sp) * abs(tan(slope) + bSlope))
}

sim = data.frame(
  slope = c(
    runif(5000, -0.8,0.8),
    rnorm(6000, 0,0.2)
  ))
sim$week = runif(length(sim$slope), 1, 10)

  
sim <- sim %>%
  mutate(
    speed = walkspeed(
      slope = slope,
      V_m = 3.4, 
      bSlope = 0.05,
      #phi = 0.99,
      week = week
    ) + rnorm(n(), 0, sd = 0.05),
    speed = ifelse(speed <= 0, NA, speed)
  ) %>%
  tbl_df()


ggplot(
  data = sim,
  aes(
    x = slope,
    y = speed,
    colour = week
  )
) + 
  geom_point() +
  theme_minimal() +
  xlab("Slope (in degrees)") +
  ylab("Walking Speed (mph)")




```


Note that at lower `week`s (experience) the hiker travels more slowly. 


## Estimating the Trail Speed Function

```{r, echo = T}

est_hs <- function(data = data){
  # data is a data.frame with: 
  # - speed var: of speeds
  # - slope var: of slopes associated with those speeds
  # - week  var: the number of weeks of hiking, experience/fitness
  
  # stage 1
  # - est Sp coefficient 
  reg.s1 <- lm(
    log(speed) ~ tan(abs(slope)),
    data = data
  ) 
  
  data$yhat <- exp(predict(reg.s1, data)) 
  data <- data %>%
    mutate(
      #truspeed = (V_m - 1.5/week) * exp(exp(-3.5 * abs(tan(slope) + 0.05)),
      resid = speed/(yhat /  exp(reg.s1$coefficients[1])),
      nuweek = 1/week
    )
  
  # stage 2 
  # - est phi fitness coefficient
  # - est V_m max speed coefficient 
  reg.s2 <- lm(
    resid ~ I(-1*nuweek),
    data = data
  )

  result <- list(
    V_m = as.numeric(reg.s2$coefficients[1]), 
    Sp =  as.numeric(reg.s1$coefficients[2]),
    phi = as.numeric(reg.s2$coefficients[2])
  )
  
  
}


est <- est_hs(sim)
est

```

```{r, cache = F, echo = F}

sliderInput("week", "Weeks of hiking:", min = 1, max = 10, value = 9)

p <-  ggplot(
  data = sim,
  aes(
    x = slope,
    y = speed,
    colour = week
  )
) + 
  geom_point(alpha = 0.5) +
  theme_minimal() +
  xlab("Slope (in degrees)") +
  ylab("Walking Speed (mph)")

renderPlot({
  
  sim.est <- data.frame(
    slope = seq(-0.7,0.7,by = 0.01),
    week = input$week
  ) %>% 
    mutate(
      speed = walkspeed(
        slope = slope, 
        Sp = est$Sp,
        V_m = est$V_m,
        phi = est$phi,
        week = week
      )
    )
  
  
  p <- p + 
    geom_line(
      data = sim.est,
      aes(
        x = slope,
        y = speed
      ),
      color = "black",
      size = 2
    ) 
  p 
  
})
```




